Fluid-mechanical erosion creates anisotropic porous media

### 2022-05-06 docs/PhysicsPaper/StokesErosion.tex
NM

Consider first an isotropic medium with {\em scalar} permeability $k$. Equivalently, we can introduce a {\em resistivity} simply defined as the reciprocal of permeability, $r = 1/k$. Then taking the horizontal component gives
\begin{equation}
\label{q1}
q_1(x,y) = -\frac{1}{r} \pderiv{p}{x}(x,y)
\end{equation}
Consider the horizontal discharge, $q_1(x,y)$, averaged over a vertical slice
\begin{equation}
\qavg := \frac{1}{2} \int_{-1}^{1} q_1(x_0, y) dy
\end{equation}
By conservation of mass, $\qavg$ is independent of the horizontal position $x_0$. Since $\bq$ is a volume-average of the tracer velocity, $\qavg$ can be equivalently obtained by averaging the tracer velocity, which is how we will measure this quantity in the simulations.

Next, consider the pressure averaged over a vertical slice
\begin{equation}
\pavg(x_0) := \frac{1}{2} \int_{-1}^{1} p(x_0, y) dy
\end{equation}
Integrating \eqref{q1} over the domain $[-x_0, x_0] \times [-1,1]$, applying the fundamental theorem of calculus, and rearranging gives
\begin{equation}
r = \left( \frac{\pavg(-x_0) - \pavg(x_0)}{2 x_0 \qavg} \right)
\end{equation}
This formula gives the resistivity based on measurements of the pressure and discharge.

>>>>

Consider first an isotropic porous medium with {\em scalar} permeability $k$. Furthermore, we seek to characterize the medium by a single, bulk value permeability, so we assume that $k$ is spatially constant.
Equivalently, we can introduce a {\em resistivity} simply defined as the reciprocal of permeability, $r = 1/k$. Then taking the horizontal component gives
\begin{equation}
\label{q1}
q_1(x,y) = -\frac{1}{r} \pderiv{p}{x}(x,y)
\end{equation}
Consider the horizontal discharge, $q_1(x,y)$, averaged over a vertical slice
\begin{equation}
\qavg := \frac{1}{2} \int_{-1}^{1} q_1(x_0, y) dy
\end{equation}
By conservation of mass, $\qavg$ is independent of the horizontal position $x_0$. Since $\bq$ is a volume-average of the tracer velocity, $\qavg$ can be equivalently obtained by averaging the tracer velocity, which is how we will measure this quantity in the simulations.

Next, consider the pressure averaged over a vertical slice
\begin{equation}
\pavg(x_0) := \frac{1}{2} \int_{-1}^{1} p(x_0, y) dy
\end{equation}
Integrating \eqref{q1} over the domain $[-x_0, x_0] \times [-1,1]$, applying the fundamental theorem of calculus, and rearranging gives
\begin{equation}
r = \left( \frac{\pavg(-x_0) - \pavg(x_0)}{2 x_0 \qavg} \right)
\end{equation}
This formula gives the resistivity based on measurements of the pressure and discharge. In particular, to compute the resistivity of the porous matrix, we must set $x_0 =1$ in principle. (If a larger value of $x_0$ is chosen, then one would be in effect averaging the porous-medium resistivity with the vanishing resistivity of the empty buffer region.)
